---
title: "Spills Workstation"
author: "allycat"
date: "`r Sys.Date()`"
output: html_document
---

This R Markdown file provides information on the base functions used in the files spill_app_daily.R and spill_app_inst.R
The functions were altered within the apps to accommodate necessary changes.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's get to it!

```{r Important libraries, message = FALSE}
library(shiny)
library(shinydashboard)
library(DT)
library(dataRetrieval)
library(dplyr)
```


### USGS Online Data Inputs

You can pull data from up to 1 day before with this code, not day of (will be old data). readNWISdv (dv = daily value)
```{r Daily mean}
# NORTH BRANCH POTOMAC RIVER AT STEYER, MD
siteNumber <- "01595000"
parameterCd <- "00060" # Discharge
startDate <- "2023-07-17"
endDate <- "2023-07-17"
discharge <- readNWISdv(siteNumber, parameterCd, startDate, endDate)
discharge <- renameNWISColumns(discharge)
discharge$X_00060_00003
```



You can pull instantaneous data at the intervals of data capture specific to the gage. readNWISuv.
```{r Unit data current}
# NORTH BRANCH POTOMAC RIVER AT STEYER, MD
siteNumber <- "01595000"
parameterCd <- "00060"
startDate1 <- "2023-07-21"
endDate1 <- "" # Leaving this blank gives most recent value
discharge1 <- readNWISuv(siteNumber, parameterCd, startDate1, endDate1, tz = "America/New_York")
discharge1 <- renameNWISColumns(discharge1)
discharge1
```



You can pull instantaneous data of a specific number of intervals. readNWISuv.
```{r Unit data current, 5 intervals}
# NORTH BRANCH POTOMAC RIVER AT STEYER, MD
discharge2 <- readNWISuv(siteNumber, parameterCd, startDate1, endDate1, tz = "America/New_York")
discharge2 <- discharge2[order(discharge2$dateTime, decreasing = TRUE), ]
num_intervals <- 5
discharge2 <- discharge2[1:num_intervals, ]
discharge2 <- renameNWISColumns(discharge2)
discharge2
discharge2$Flow_Inst
```
*order() lets you sort data by dateTime in descending order to get most recent intervals first.*



You can calculate velocity using most recent discharge rate(s).
```{r Unit data current, velocity}
# NORTH BRANCH POTOMAC RIVER AT STEYER, MD
a <- 0.2 # Dummy 
b <- 0.3 # Dummy

flow_values <- discharge2$Flow_Inst
flow_value_now <- discharge2$Flow_Inst[1]

velocity_5 <- a*flow_values^b
velocity_5 # 5 intervals

velocity_now <- a*flow_value_now^b
velocity_now # most recent value only
```
*Velocity equation is V = a times Q^b, with Q being flow rate, and a/b being coefficients specific to each stream.*


```{r Daily mean}
siteNumber <- "01595000"
parameterCd <- "00060" # Discharge
startDate <- "2023-07-21"
endDate <- "2023-07-21"
discharge <- readNWISdv(siteNumber, parameterCd, startDate, endDate)
discharge1 <- discharge$X_00060_00003
discharge1
```
